---
title: "07_lagged_models"
author: "Levente Ronai"
output: html_document
  html_document:
   theme: spacelab
   code_download: true
   toc: true
   toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r, include=F}

options(scipen = 999)

suppressPackageStartupMessages(library(ppcor))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(rstatix))
suppressPackageStartupMessages(library(xts))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(library(qgraph))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(lmerTest))
suppressPackageStartupMessages(library(sjPlot))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(broom))
suppressPackageStartupMessages(library(performance))
suppressPackageStartupMessages(library(lmeresampler))

```

import data

```{r, message=F, warning=F}
dat_cleaned <- read_csv("data_processed/dat_cleaned.csv")

model_dat <- read.csv("data_processed/model_dat.csv")
```
# lagging ples, event appraisal and stressor-exposure

```{r}
data_lagged <- 
  model_dat %>% 
  dplyr::select(id_esm, age, sex = gender, day, beep, diary_counter, event_unpleasantness,
                event_unpleasantness_centered, ple_cfa, stress_exposure_centered = stressor_mean_centered_int, stress_exposure = stressor_mean, event_unpleasantness_mean = event_unpleasantness_3days_mean, contains("mss")) %>% 
  group_by(id_esm, diary_counter) %>%
  mutate(
    event_unpleasantness_centered_mean = mean(event_unpleasantness_centered, na.rm = T),
    ple_cfa_mean = mean(ple_cfa, na.rm = T)) %>% 
  ungroup() %>% 
  group_by(id_esm) %>%
  mutate(
    ple_cfa_3days_lag = lag(ple_cfa, 72),
    event_unpleasantness_centered_3days_lag = lag(event_unpleasantness_centered, 72),
    stress_exposure_centered_lag = lag(stress_exposure_centered, 24)) %>% 
  ungroup() %>% 
  group_by(id_esm, day) %>%
  mutate(
    ple_cfa_lag = lag(ple_cfa),
    event_unpleasantness_centered_lag = lag(event_unpleasantness_centered)) %>% 
  ungroup() 
  
```

# lagged relationships between the above variables in different settings across 3-day measurement windows (diary_counter)

```{r}
lagged_1 <- 
  lmer(ple_cfa ~ 
         sex + age + day + beep + diary_counter + 
         stress_exposure_centered_lag +
         (stress_exposure_centered_lag | id_esm), data = data_lagged)

lagged_2 <- 
  lmer(stress_exposure ~ 
         sex + age + day + beep + diary_counter +
         ple_cfa_3days_lag + 
         (ple_cfa_3days_lag  | id_esm), data = data_lagged)


lagged_3 <- 
  lmer(event_unpleasantness ~ 
         sex + age + day + beep + diary_counter +
         stress_exposure_centered_lag +
         (stress_exposure_centered_lag | id_esm), data = data_lagged)

lagged_4 <- 
  lmer(stress_exposure ~ 
         sex + age + diary_counter + event_unpleasantness_centered_3days_lag + 
         (event_unpleasantness_centered_3days_lag  | id_esm), data = data_lagged)


tab_model(lagged_1,
          lagged_3, 
          lagged_2, 
          lagged_4, 
          show.std = T)

```

# testing possible confounder effect of stressor exposure on the relationship between event unpleasantness and pleasure within individuals using partial correlations in the 3-day windows

```{r}
model_within_corr <- 
  model_dat %>% 
  dplyr::select(id_esm, age, gender, day, beep, diary_counter,
                event_unpleasantness_centered, ple_cfa, stressor_mean_centered_int) %>% 
  distinct() %>%
  group_by(id_esm, diary_counter) %>%
  summarise(
    event_unpleasantness_centered_mean = mean(event_unpleasantness_centered, na.rm = T),
    ple_cfa_mean = mean(ple_cfa, na.rm = T),
    stressor_mean_centered_int_mean = mean(stressor_mean_centered_int, na.rm = T)) %>% 
  ungroup() %>% 
  filter_at(vars(event_unpleasantness_centered_mean, ple_cfa_mean, stressor_mean_centered_int_mean), all_vars(!is.na(.))) %>% 
  group_by(id_esm) %>%
  filter(n_distinct(diary_counter) > 2) %>%
  ungroup()
  
```

```{r, warning=F}
# event_unpleasantness_centered_mean
# ple_cfa_mean
# stressor_mean_centered_int_mean

# Apply the function to each individual
results_partial <- 
  model_within_corr %>%
  group_by(id_esm) %>%
  summarise(coef = pcor.test(event_unpleasantness_centered_mean, 
                                         ple_cfa_mean, stressor_mean_centered_int_mean, method = "spearman")$estimate) %>% 
  filter(!is.na(coef)) %>% 
  mutate(partial = "stressor_as_confounder") 

results_partial_2 <- 
  model_within_corr %>%
  group_by(id_esm) %>%
  summarise(coef = pcor.test(event_unpleasantness_centered_mean, stressor_mean_centered_int_mean,
                                         ple_cfa_mean, method = "spearman")$estimate) %>% 
  filter(!is.na(coef)) %>% 
  mutate(partial = "ple_as_confounder") 

results_partial_3 <- 
  model_within_corr %>%
  group_by(id_esm) %>%
  summarise(coef = pcor.test(stressor_mean_centered_int_mean,
                                         ple_cfa_mean, event_unpleasantness_centered_mean,method = "spearman")$estimate) %>% 
  filter(!is.na(coef)) %>% 
  mutate(partial = "event_unpl_as_confounder") 


results <- 
  model_within_corr %>%
  group_by(id_esm) %>%
  summarise(coef = cor.test(event_unpleasantness_centered_mean, 
                                         ple_cfa_mean, method = "spearman")$estimate) %>% 
  filter(!is.na(coef)) %>% 
  mutate(partial = "event_ple_corr") 

results_2 <- 
  model_within_corr %>%
  group_by(id_esm) %>%
  summarise(coef = cor.test(stressor_mean_centered_int_mean, 
                                         ple_cfa_mean, method = "spearman")$estimate) %>% 
  filter(!is.na(coef)) %>% 
  mutate(partial = "stressor_ple_corr") 

results_3 <- 
  model_within_corr %>%
  group_by(id_esm) %>%
  summarise(coef = cor.test(stressor_mean_centered_int_mean, 
                                         event_unpleasantness_centered_mean, method = "spearman")$estimate) %>% 
  filter(!is.na(coef)) %>% 
  mutate(partial = "stressor_event_corr") 

comb <- 
  results_partial %>% 
  rbind(results, results_partial_2, results_partial_3, results_2, results_3)

```


```{r}
# descriptives for partial_corrs$partial_corr
comb %>% 
  group_by(partial) %>% 
  summarise(mean = round(mean(coef), 2),
            meadian = round(median(coef),2),
            sd = round(sd(coef), 2)) %>% 
  filter(partial == "event_ple_corr" | partial == "stressor_as_confounder")
```
```{r}
comb %>% 
  group_by(partial) %>% 
  summarise(mean = round(mean(coef), 2),
            meadian = round(median(coef),2),
            sd = round(sd(coef), 2)) %>% 
  filter(partial == "stressor_event_corr" | partial == "ple_as_confounder")
```


```{r}
comb %>% 
  group_by(partial) %>% 
  summarise(mean = round(mean(coef), 2),
            meadian = round(median(coef),2),
            sd = round(sd(coef), 2)) %>% 
  filter(partial == "stressor_ple_corr" | partial == "event_unpl_as_confounder")
```
# histograms per partial corr coefficents 

```{r}
comb %>% 
  ggplot() +
  aes(x = coef, fill = partial) +
  geom_histogram(binwidth = 0.15) +
  facet_wrap(~partial) +
  theme_minimal() +
  # black line to medians
```


